---
title: "Simulation: Recovering ACT-R parameters using the LBA"
author: "Maarten van der Velde"
date: "Last updated: `r Sys.Date()`"
output:
  html_notebook:
    smart: no
    toc: yes
    toc_float: yes
  github_document:
    toc: yes
editor_options: 
  chunk_output_type: inline
---



# Overview

In this notebook, we generate retrieval data using an ACT-R model, and then attempt to recover the original ACT-R parameters using an LBA.

We perform several simulations (only 3 and 4 are reported in the paper):

1. **Single participant doing 100,000 trials.**<br>
To get an idea of best-case recovery performance.
2. **Single participant doing a varying number of trials.**<br>
To see how recovery performance changes with the size of the available data.
3. **Multiple participants doing 100 trials.**<br>
To see how well individual parameters can be recovered in a "typical" experimental sample.
4. **Multiple participants doing a varying number of trials.**<br>
To see how individual parameter recovery performance changes with the size of the available data.



# Setup
```{r}
library(dplyr)
library(ggplot2)
library(rtdists)
library(purrr)
library(furrr)
library(tidyr)
library(truncdist)
library(cowplot)
library(grid)
library(rlang)

future::plan("multiprocess", workers = 6) # Set to desired number of cores

theme_paper <- theme_classic(base_size = 14) + 
  theme(axis.text = element_text(colour = "black"))


use_cached_results <- TRUE # Set to FALSE to rerun simulations (takes time!)
```

## Parameter recovery functions

Define the objective function to minimise (adapted to lognormal distribution from the [rtdists example](https://github.com/rtdists/rtdists/blob/master/examples/examples.lba.R)):
```{r}
obj_fun <- function(par, rt, response, distribution = "lnorm") {
  # simple parameters
  spar <- par[!grepl("[12]$", names(par))]  
  
  # distribution parameters:
  dist_par_names <- unique(sub("[12]$", "", grep("[12]$" ,names(par), value = TRUE)))
  dist_par <- vector("list", length = length(dist_par_names))
  names(dist_par) <- dist_par_names
  for (i in dist_par_names) dist_par[[i]] <- as.list(unname(par[grep(i, names(par))]))
  dist_par$sdlog_v <- c(1, dist_par$sdlog_v) # fix first sdlog_v to 1

  # get summed log-likelihood:
  d <- do.call(dLBA, args = c(rt=list(rt), response=list(response), spar, dist_par, 
                               distribution=distribution, silent=TRUE))
  if (any(d < 0e-10)) return(1e6) 
  else return(-sum(log(d)))
}
```

Define the parameter recovery function, which randomly initialises the LBA parameters (within some reasonable constraints to promote convergence) and then uses the nlminb optimiser to find the best fit, evaluating each iteration with the objective function:
```{r}
recover_parameters <- function(data, obj_fun) {
  
  # Generate random starting values
  init_par <- runif(6)
  init_par[2] <- init_par[2] + 1 # Ensure b is larger than A
  init_par[3] <- runif(1, 0, min(data$rt)) # Ensure t0 is mot too large
  init_par[4] <- -init_par[4] # Ensure meanlog_v1 is negative
  init_par[5] <- init_par[4] - init_par[5] # Ensure meanlog_v2 is negative and lower than meanlog_v2
  names(init_par) <- c("A", "b", "t0", "meanlog_v1", "meanlog_v2", "sdlog_v2")
  
  # Run optimiser
  fit <- nlminb(obj_fun,
                start = init_par, 
                rt = data$rt, response = data$response,
                lower = c(0, 0, 0, -Inf, -Inf, 0)) # Set lower bounds on parameters
  
  # Only keep parameter estimates if the optimiser converged successfully
  if (fit$convergence == 0 && !is.infinite(fit$objective)) {
    return(as.list(c(fit$par, objective = fit$objective)))
  }
  
  return(NULL)
  
}
```



# Simulation 1: single participant, 100,000 trials

We'll use an ACT-R model of a single participant to generate response data (RT and choice) for a retrieval task with two response options.
In the context of cued retrieval, we can think of these two options as representing (1) the correct response and (2) all possible incorrect responses.

The simulated participant will have a single latency factor $F$ and non-retrieval time $t_{er}$, but the activation of the memory chunks will be drawn from a normal distribution to represent the activation noise component in ACT-R.

## Generate data

```{r}
set.seed(2021)
```


Simulate 100,000 trials:
```{r}
n_trials <- 1e5
```

Set the ACT-R parameters:
```{r}
# Latency factor F
lf <- 2

# Non-retrieval time t_er
t_er <- .5

# Activation of correct answer
a_c_mu <- -.5
a_c_sd <- 1

# Activation of incorrect answer
a_f_mu <- -1.5
a_f_sd <- 1.5
```

Generate the data:
```{r}
sim_actr <- tibble(
  f = rep(lf, n_trials),
  a_c = rnorm(n_trials, mean = a_c_mu, sd = a_c_sd),
  a_f = rnorm(n_trials, mean = a_f_mu, sd = a_f_sd),
  t = rep(t_er, n_trials)
) %>%
  rowwise() %>%
  mutate(rt = f * exp(-max(a_c, a_f)) + t,
         response = ifelse(a_c > a_f, 1, 2)) %>%
  ungroup()
```


```{r}
head(sim_actr)
prop.table(table(sim_actr$response))
```

Note: a response of 1 indicates a correct answer, 2 an incorrect answer.


To confirm that the data generation process of ACT-R is equivalent to that of the LBA, we'll also calculate the density function of an LBA with the same parameters.

The `rtdists` package allows us to define drift rate with a lognormal distribution, which means that we can keep the untransformed (i.e., log-scale) activation mean and SD from ACT-R.

```{r}
# F = b - A/2 -> b = F + A/2
# Set A to 1 to find the value of b
A <- 1
b <- lf + .5*A

dlba_dat <- crossing(rt = seq(0, 20, by = .01),
                     response = c(1, 2))

sim_lba <- bind_cols(dlba_dat,
                     density = dLBA(rt = dlba_dat$rt,
                                    response = dlba_dat$response,
                                    A = A,
                                    b = b,
                                    t0 = t_er,
                                    meanlog_v = c(a_c_mu, a_f_mu),
                                    sdlog_v = c(a_c_sd, a_f_sd),
                                    distribution = "lnorm",
                                    silent = TRUE)) %>%
  mutate(rt = ifelse(response == 1, rt, -rt),
         model = "LBA")
```



As the plot below shows, there is an excellent correspondence between the distribution of responses of the ACT-R model (black histogram) and the density function of the LBA (orange curve).
```{r}
sim_actr %>%
  mutate(rt = ifelse(response == 1, rt, -rt),
         model = "ACT-R") %>%
  ggplot(aes(x = rt, colour = model)) +
  geom_vline(xintercept = 0, lty = 2, colour = "grey80") +
  geom_histogram(aes(y = ..density..), binwidth = .25, colour = "black", fill = "white", size = .1) +
  geom_line(data = sim_lba, aes(y = density), size = rel(1.1), alpha = .8) +
  scale_x_continuous(limits = c(-20, 20)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = c("#e66101", "#5e3c99")) +
  labs(x = "RT (s)",
       y = "Density",
       colour = NULL) +
  guides(colour = FALSE) +
  theme_paper +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank())
```


## Recover parameters

```{r message = FALSE, warning = FALSE}
if (use_cached_results) {
  
  s1_param_recov_best <- readRDS(file.path("..", "output", "s1_param_recov_best.rds"))
  
} else {
  
  n_attempts <- 100 # More attempts = more chances to reach global optimum
  s1_param_recov <- future_map_dfr(1:n_attempts, function(x) recover_parameters(data = sim_actr, obj_fun = obj_fun))
  saveRDS(s1_param_recov, file.path("..", "output", "s1_param_recov.rds"))

  # Only keep the best fit across all attempts, which is closest to the global optimum
  s1_param_recov_best <- filter(s1_param_recov, objective == min(objective))
  saveRDS(s1_param_recov_best, file.path("..", "output", "s1_param_recov_best.rds"))

}
```


To compare the recovered parameter estimates to the original ACT-R parameters that generated the data, transform from LBA parameters to ACT-R parameters:
```{r}
f_recov <- s1_param_recov_best$b - .5*s1_param_recov_best$A
t_recov <- s1_param_recov_best$t0
a_c_recov <- s1_param_recov_best$meanlog_v1
a_f_recov <- s1_param_recov_best$meanlog_v2
a_c_sd_recov <- rep(1, nrow(s1_param_recov_best))
a_f_sd_recov <- s1_param_recov_best$sdlog_v2


s1_par_comp <- tibble(parameter = rep(c("F", "t_er", "A_correct", "A_incorrect", "A_correct_sd", "A_incorrect_sd"), each = nrow(s1_param_recov_best)),
                      original = rep(c(lf, t_er, a_c_mu, a_f_mu, a_c_sd, a_f_sd), each = nrow(s1_param_recov_best)),
                      recovered = c(f_recov, t_recov, a_c_recov, a_f_recov, a_c_sd_recov, a_f_sd_recov))
```

Comparison of original ACT-R parameters to recovered parameter estimates from the LBA:
```{r}
ggplot(s1_par_comp, aes(x = original, y = recovered, colour = parameter)) +
  geom_abline(lty = 2) +
  geom_point() +
  coord_equal() +
  labs(x = "Original parameter value",
       y = "Recovered parameter value",
       colour = "Parameter") +
  theme_paper
```

Difference between the original ACT-R parameters and the recovered parameter estimates from the LBA:
```{r}
ggplot(s1_par_comp, aes(x = parameter, y = original - recovered)) +
  geom_hline(yintercept = 0, lty = 2) +
  geom_point(aes(colour = parameter)) +
  ylim(-max(abs(s1_par_comp$original - s1_par_comp$recovered)), max(abs(s1_par_comp$original - s1_par_comp$recovered))) +
  guides(colour = FALSE) +
  labs(x = NULL,
       y = "Original - recovered parameter value") +
  theme_paper
```




# Simulation 2: single participant, vary dataset size

How accurately can we recover parameters from smaller datasets?

```{r message = FALSE, warning = FALSE}
if (use_cached_results) {
  
  s2_param_recov_best <- readRDS(file.path("..", "output", "s2_param_recov_best.rds"))
  
} else {
  
  dataset_sizes <-  c(25, 50, 100, 250, 500, 1000, 2500, 5000, 10000, 25000, 50000)
  n_attempts <- 100
  s2_param_recov <- list()
  
  for(i in 1:length(dataset_sizes)) {
    s2_param_recov[[i]] <- future_map_dfr(1:n_attempts, function(x) recover_parameters(data = sim_actr[1:dataset_sizes[i],], obj_fun = obj_fun)) %>%
      mutate(dataset_size = dataset_sizes[i])
    
  }
  
  s2_param_recov <- bind_rows(s2_param_recov)
  saveRDS(s2_param_recov, file.path("..", "output", "s2_param_recov.rds"))
  
  # Only keep the best fit across all attempts, which is closest to the global optimum
  s2_param_recov_best <- s2_param_recov %>%
    group_by(dataset_size) %>%
    filter(objective == min(objective)) %>%
    slice(n()) %>% # If there are multiple rows with the exact same value, only keep the first
    ungroup()
  
  s2_param_recov_best <- bind_rows(s2_param_recov_best,
                                   mutate(s1_param_recov_best, dataset_size = 100000))
  saveRDS(s2_param_recov_best, file.path("..", "output", "s2_param_recov_best.rds"))
  
}
```


To compare the recovered parameter estimates to the original ACT-R parameters that generated the data, transform from LBA parameters to ACT-R parameters:
```{r}
f_recov <- s2_param_recov_best$b - .5*s2_param_recov_best$A
t_recov <- s2_param_recov_best$t0
a_c_recov <- s2_param_recov_best$meanlog_v1
a_f_recov <- s2_param_recov_best$meanlog_v2
a_c_sd_recov <- rep(1, nrow(s2_param_recov_best))
a_f_sd_recov <- s2_param_recov_best$sdlog_v2


s2_par_comp <- tibble(parameter = rep(c("F", "t_er", "A_correct", "A_incorrect", "A_correct_sd", "A_incorrect_sd"), each = nrow(s2_param_recov_best)),
                   original = rep(c(lf, t_er, a_c_mu, a_f_mu, a_c_sd, a_f_sd), each = nrow(s2_param_recov_best)),
                   recovered = c(f_recov, t_recov, a_c_recov, a_f_recov, a_c_sd_recov, a_f_sd_recov),
                   dataset_size = rep(s2_param_recov_best$dataset_size, 6))
```

Comparison of original ACT-R parameters to recovered parameter estimates from the LBA:
```{r fig.width = 9}
ggplot(mutate(s2_par_comp, n = dataset_size), aes(x = original, y = recovered, colour = parameter)) +
  facet_wrap(~ n, labeller = function(x) label_both(labels = x, sep = " = ")) +
  geom_abline(lty = 2) +
  geom_point() +
  coord_equal() +
  labs(x = "Original parameter value",
         y = "Recovered parameter value",
         colour = "Parameter") +
  theme_paper
```

Difference between the original ACT-R parameters and the recovered parameter estimates from the LBA, as a function of datset size:
```{r}
s2_par_comp %>%
  rowwise() %>%
  mutate(ae = abs(original - recovered)) %>%
  group_by(dataset_size) %>%
  mutate(mae = mean(ae)) %>%
  ggplot(aes(x = dataset_size, y = mae)) +
  geom_line(lty = 2) +
  geom_jitter(aes(y = ae, colour = parameter), height = 0, width = .025, alpha = .75) +
  geom_point(size = rel(1.5)) +
  scale_x_log10(labels = scales::comma_format()) +
  labs(x = "Dataset size",
       y = "Absolute error") +
  guides(colour = FALSE) +
  theme_paper 
```



# Simulation 3: multiple participants, 100 trials

Here, we extend the previous example by simulating data from multiple participants, each with their own memory parameters.

Each participant will have a latency factor $F \sim N_+(1, .5)$  and a non-retrieval time $t_{er} \sim N_+(.75, .5)$.
The activation of their memory chunks will be drawn from two normal distributions, with mean $\mu_{c} \sim N_-(-.5, .5)$ and $\mu_{f} \sim N_-(-1.5, .5)$, and standard deviation $\sigma_{c} = 1$ and $\sigma_{f} \sim N_+(1.5, .5)$.

## Generate data

```{r}
set.seed(2021)
```

```{r}
# Number of trials to simulate per participant:
n_trials <- 100

# Number of participants to simulate:
n_participants <- 25
```

Set the ACT-R parameters:
```{r}
# Latency factor F
lf <- rtrunc(n_participants, spec = "norm", mean = 1, sd = .5, a = 0, b = Inf)

# Non-retrieval time t_er
t_er <- rtrunc(n_participants, spec = "norm", mean = .75, sd = .5, a = 0, b = Inf)

# Activation of correct answer
a_c_mu <- rtrunc(n_participants, spec = "norm", mean = -.5, sd = .5, a = -Inf, b = 0)
a_c_sd <- rep(1, n_participants)

# Activation of incorrect answer
a_f_mu <- rtrunc(n_participants, spec = "norm", mean = -1.5, sd = .5, a = -Inf, b = 0)
a_f_sd <- rtrunc(n_participants, spec = "norm", mean = 1.5, sd = .5, a = 0, b = Inf)
```

Distribution of parameters:
```{r}
tibble(participant = 1:n_participants, lf, t_er, a_c_mu, a_c_sd, a_f_mu, a_f_sd) %>%
  pivot_longer(-participant, "parameter") %>%
  ggplot(aes(x = parameter, y = value, colour = parameter)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(height = 0, width = .1) +
  labs(x = "Parameter",
         y = "Value",
         colour = "Parameter") +
  guides(colour = FALSE) +
  theme_paper
```

Alternative visualisation:
```{r}
tibble(participant = 1:n_participants, lf, t_er, a_c_mu, a_c_sd, a_f_mu, a_f_sd) %>%
  pivot_longer(-participant, "parameter") %>%
  mutate(parameter = factor(parameter, 
                            levels = c("a_c_mu", "a_f_mu", "a_c_sd", "a_f_sd", "lf", "t_er"),
                            labels  = c(expression(mu[c]), expression(mu[f]), expression(sigma[c]), expression(sigma[f]), expression(F), expression(t[er])))) %>%
  ggplot(aes(y = parameter, x = value, colour = (parameter))) +
  facet_grid(parameter ~ ., scales = "free", switch = "y", labeller = labeller(parameter = label_parsed)) +
  # geom_boxplot(outlier.shape = NA) +
  geom_hline(aes(yintercept = parameter), colour = "grey90", lty = 3) +
  geom_jitter(height = .1, width = 0, size = .5) +
  labs(x = NULL,
       y = NULL) +
  guides(colour = FALSE) +
  scale_colour_viridis_d() +
  theme_paper +
  theme(strip.background = element_blank(),
        strip.text.y.left = element_text(angle = 0, size = rel(1.25)),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank())

ggsave(file.path("..", "output", "sim-param-values.png"), width = 2, height = 2, dpi = 600)
```

Generate the data:
```{r}
sim_actr <- tibble(participant = rep(1:n_participants, each = n_trials),
       f = rep(lf, each = n_trials),
       a_c = rnorm(n_participants * n_trials, mean = rep(a_c_mu, each = n_trials), sd = rep(a_c_sd, each = n_trials)),
       a_f = rnorm(n_participants * n_trials, mean = rep(a_f_mu, each = n_trials), sd = rep(a_f_sd, each = n_trials)),
       t = rep(t_er, each = n_trials)
) %>%
  rowwise() %>%
  mutate(rt = f * exp(-max(a_c, a_f)) + t,
         response = ifelse(a_c > a_f, 1, 2)) %>%
  ungroup()
```

```{r}
head(sim_actr)
prop.table(table(sim_actr$response))
```

RT distributions:
```{r}
ggplot(mutate(sim_actr, rt = ifelse(response == 1, rt, -rt)), aes(x = rt, group = participant, colour = participant)) +
  geom_line(stat = "density", adjust = .5, n = 2^12, na.rm = F, alpha = .5) +
  geom_vline(aes(xintercept = 0), lty = 2, colour = "grey80") +
  scale_x_continuous(breaks = c(-15, 0, 15)) +
  scale_y_continuous(expand = c(0,0), breaks = NULL) +
  coord_cartesian(xlim = c(-20, 20)) +
  labs(x = "RT (s)",
       y = "Density") +
  guides(colour = FALSE) +
  theme_paper +
  theme(strip.background = element_blank(),
        strip.text = element_blank())
```


## Recover parameters

```{r message = FALSE, warning = FALSE}
if (use_cached_results) {
  
  s3_param_recov_best <- readRDS(file.path("..", "output", "s3_param_recov_best.rds"))
  
} else {
  
  n_attempts <- 250 # More attempts = more chances to reach global optimum
  
  sim_actr_split <- split(sim_actr, sim_actr$participant)
  
  s3_param_recov <- future_map_dfr(sim_actr_split, function(x) {
    future_map_dfr(1:n_attempts, function(y) {
      recover_parameters(data = x, obj_fun = obj_fun)
    }) %>%
      mutate(participant = x$participant[1])
  })
  
  saveRDS(s3_param_recov, file.path("..", "output", "s3_param_recov.rds"))
  
  # Only keep the best fit across all attempts
  s3_param_recov_best <- s3_param_recov %>%
    group_by(participant) %>%
    filter(objective == min(objective)) %>%
    slice(1) # If there are multiple equally good fits, just keep the first
  
  saveRDS(s3_param_recov_best, file.path("..", "output", "s3_param_recov_best.rds"))
  
}
```


```{r}
f_recov <- s3_param_recov_best$b - .5*s3_param_recov_best$A
t_recov <- s3_param_recov_best$t0
a_c_recov <- s3_param_recov_best$meanlog_v1
a_f_recov <- s3_param_recov_best$meanlog_v2
a_c_sd_recov <- rep(1, n_participants)
a_f_sd_recov <- s3_param_recov_best$sdlog_v2


s3_par_comp <- tibble(
  participant = rep(1:n_participants, 6),
  parameter = rep(c("F", "t_er", "A_correct", "A_incorrect", "A_correct_sd", "A_incorrect_sd"), each = n_participants),
  original = c(lf, t_er, a_c_mu, a_f_mu, a_c_sd, a_f_sd),
  recovered = c(f_recov, t_recov, a_c_recov, a_f_recov, a_c_sd_recov, a_f_sd_recov)
)
```

Comparison of original ACT-R parameters to recovered parameter estimates from the LBA:
```{r}
s3_par_comp_plot <- s3_par_comp %>%
  filter(parameter != "A_correct_sd") %>%
  mutate(parameter = factor(parameter, 
                            levels = c("A_correct", "A_incorrect", "A_incorrect_sd", "F", "t_er"),
                            labels  = c(expression(mu[c]), expression(mu[f]), expression(sigma[f]), expression(F), expression(t[er]))))

p_par_comp <- ggplot(s3_par_comp_plot, aes(x = original, y = recovered)) +
  facet_wrap(~ parameter, ncol = 3, scales = "free_x", labeller = labeller(parameter = label_parsed))+
  geom_abline(lty = 2, colour = "grey80") +
  geom_smooth(method = "lm", formula = y ~ x, alpha = .5, colour = "black", lwd = rel(.8)) +
  geom_point(aes(colour = parameter), alpha = .5) +
  # coord_fixed() +
  labs(x = "Original",
         y = "Recovered",
         colour = NULL) +
  scale_x_continuous(limits = c(min(s3_par_comp$recovered), max(s3_par_comp$recovered))) +
  scale_colour_viridis_d() +
  guides(colour = FALSE) +
  theme_paper +
  theme_paper +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = rel(1)))

p_par_comp
```

Alternative plot:
```{r}
ggplot(s3_par_comp_plot, aes(x = original, y = recovered)) +
  facet_wrap(~ parameter, ncol = 5, labeller = labeller(parameter = label_parsed))+
  geom_abline(lty = 2, colour = "grey80") +
  geom_smooth(method = "lm", formula = y ~ x, alpha = .5, colour = "black", lwd = rel(.8)) +
  geom_point(aes(colour = parameter), alpha = .5) +
  coord_fixed() +
  labs(x = "Original",
         y = "Recovered",
         colour = NULL) +
  scale_x_continuous(limits = c(min(s3_par_comp$recovered), max(s3_par_comp$recovered))) +
  scale_colour_viridis_d() +
  guides(colour = FALSE) +
  theme_paper +
  theme_paper +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = rel(1)))

ggsave(file.path("..", "output", "param-recov-comparison.png"), width = 6.5, height = 3, dpi = 600)
```



Compare fitted LBA density to data:
```{r}
dlba_dat <- crossing(rt = seq(0, 20, by = .01),
         response = c(1, 2))

sim_lba <- s3_param_recov_best %>%
  split(.$participant) %>%
  future_map_dfr(function (x) {
    bind_cols(dlba_dat, density =
    dLBA(rt = dlba_dat$rt,
         response = dlba_dat$response,
         A = x$A,
         b = x$b,
         t0 = x$t0,
         meanlog_v = c(x$meanlog_v1, x$meanlog_v2),
         sdlog_v = c(1, x$sdlog_v2),
         distribution = "lnorm",
         silent = TRUE)) %>%
      mutate(participant = x$participant[1])
  }) %>%
  mutate(rt = ifelse(response == 1, rt, -rt),
         model = "LBA")
```

```{r fig.width = 9}
sim_actr %>%
  mutate(rt = ifelse(response == 1, rt, -rt),
         model = "Data") %>%
  ggplot(aes(x = rt, colour = model)) +
  facet_wrap(~ participant, ncol = 5) +
  geom_vline(xintercept = 0, lty = 2, colour = "grey80") +
  geom_histogram(aes(y = ..density..), binwidth = .5, colour = "black", fill = "white", size = .1) +
  geom_line(data = sim_lba, aes(y = density)) +
  scale_x_continuous(limits = c(-20, 20), breaks = c(-15, 0, 15)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = c("#e66101", "#5e3c99")) +
  labs(x = "RT (s)",
       y = "Density",
       colour = NULL) +
  guides(colour = FALSE) +
  theme_paper +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank())
```


Alternative plot for presentation:
```{r}

draw_key_custom <- function(data, params, size) {
  if(data$colour == "#000000" && data$size == .1) { # ACT-R
    grobTree(
      linesGrob(
        c(.1, .1, .3, .3, .3, .5, .5, .5, .7, .7, .7, .9, .9),
        c(0, .5, .5, 0, .8, .8, 0, .65, .65, 0, .4, .4, 0)
      ),
      gp = gpar(
        col = data$colour %||% "grey20",
        fill = alpha(data$fill %||% "white", data$alpha),
        lwd = (data$size %||% 0.5) * .pt,
        lty = data$linetype %||% 1
      )
    )
  } 
  else if (data$colour == "#e66101") { # LBA
    grobTree(
      linesGrob(
        c(0, 1),
        c(.5, .5)
      ),
      gp = gpar(
        col = alpha(data$colour %||% "grey20", data$alpha),
        fill = alpha(data$fill %||% "white", data$alpha),
        lwd = (data$size %||% 0.5) * .pt,
        lty = data$linetype %||% 1
      )
    )
  }
  else {
    grobTree() # Don't draw
  }
}


sim_actr %>%
  filter(participant <= 3) %>%
  mutate(rt = ifelse(response == 1, rt, -rt),
         model = "ACT-R") %>%
  ggplot(aes(x = rt, colour = model)) +
  facet_wrap(~ participant, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, colour = "grey80") +
  geom_histogram(aes(y = ..density..), binwidth = .5, fill = "white", size = .1, key_glyph = draw_key_custom) +
  geom_line(data = filter(sim_lba, participant <= 3), aes(y = density), key_glyph = draw_key_custom) +
  scale_x_continuous(limits = c(-20, 20), breaks = c(-15, 0, 15)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_manual(values = c("#000000", "#e66101")) +
  labs(x = "RT (s)",
       y = "Density",
       colour = NULL) +
  theme_paper +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        # legend.background = element_blank(),
        legend.position = "top",
        legend.justification = "left",
        legend.direction = "vertical",
        legend.box.margin = unit(c(-20, 0, -40, -30), "pt"))

ggsave(file.path("..", "output", "param-recov-dist-comp.png"), width = 1.75, height = 3, dpi = 600)
```


# Simulation 4: multiple participants, vary dataset size

The previous simulation uses 100 trials per participant.
How accurately can we recover parameters from datasets of different sizes?

Use the same parameters as in Simulation 3:
```{r}
n_participants <- 25

set.seed(2021)

# Latency factor F
lf <- rtrunc(n_participants, spec = "norm", mean = 1, sd = .5, a = 0, b = Inf)

# Non-retrieval time t_er
t_er <- rtrunc(n_participants, spec = "norm", mean = .75, sd = .5, a = 0, b = Inf)

# Activation of correct answer
a_c_mu <- rtrunc(n_participants, spec = "norm", mean = -.5, sd = .5, a = -Inf, b = 0)
a_c_sd <- rep(1, n_participants)

# Activation of incorrect answer
a_f_mu <- rtrunc(n_participants, spec = "norm", mean = -1.5, sd = .5, a = -Inf, b = 0)
a_f_sd <- rtrunc(n_participants, spec = "norm", mean = 1.5, sd = .5, a = 0, b = Inf)
```


Define dataset sizes:
```{r}
dataset_sizes <- c(25, 50, 250, 500, 1000, 2500, 5000, 10000, 25000, 50000)
```

Generate datasets of different sizes and recover the parameters:
```{r message = FALSE, warning = FALSE}
if (use_cached_results) {
  
  s4_param_recov <- bind_rows(readRDS(file.path("..", "output", "s4_param_recov.rds")),
                              readRDS(file.path("..", "output", "s3_param_recov_best.rds")) %>%
                                mutate(dataset_size = 100))
    
} else {
  
  n_attempts <- 250
  s4_param_recov <- tibble()
  
  for(i in 1:length(dataset_sizes)) {
    
    # Generate dataset
    n_trials <- dataset_sizes[i]
    
    sim_actr <- tibble(participant = rep(1:n_participants, each = n_trials),
                       f = rep(lf, each = n_trials),
                       a_c = rnorm(n_participants * n_trials, mean = rep(a_c_mu, each = n_trials), sd = rep(a_c_sd, each = n_trials)),
                       a_f = rnorm(n_participants * n_trials, mean = rep(a_f_mu, each = n_trials), sd = rep(a_f_sd, each = n_trials)),
                       t = rep(t_er, each = n_trials)
    ) %>%
      rowwise() %>%
      mutate(rt = f * exp(-max(a_c, a_f)) + t,
             response = ifelse(a_c > a_f, 1, 2)) %>%
      ungroup()
    
    # Fit each participant separately
    sim_actr_split <- split(sim_actr, sim_actr$participant)
    
    s4_fit <- future_map_dfr(sim_actr_split, function(x) {
      
      # Run n_attempts times with different random starting values
      fit <- future_map_dfr(1:n_attempts, function(y) {
        recover_parameters(data = x, obj_fun = obj_fun)
      })
      
      # Keep the best fit
      fit %>%
        filter(objective == min(objective)) %>%
        slice(1) %>% # If there are multiple rows with the exact same value, only keep the first
        mutate(participant = x$participant[1])
      
    }) %>%
      mutate(dataset_size = dataset_sizes[i])
    
    saveRDS(s4_fit, file.path("..", "output", paste0("s4_param_recov_", dataset_sizes[i], ".rds")))
    
    s4_param_recov <- bind_rows(s4_param_recov, s4_fit)
    
  }
  
  saveRDS(s4_param_recov, file.path("..", "output", "s4_param_recov.rds"))
  
}
```


```{r}
s4_param_recov_long <- s4_param_recov %>%
  transmute(
    n = dataset_size,
    participant,
    F = b - .5*A,
    t_er = t0,
    A_correct = meanlog_v1,
    A_incorrect = meanlog_v2,
    A_correct_sd = 1,
    A_incorrect_sd = sdlog_v2
  ) %>%
  pivot_longer(F:A_incorrect_sd, names_to = "parameter", values_to = "recovered")

s4_param_orig_long <- tibble(n = rep(c(dataset_sizes, 100), each = n_participants),
                             participant = rep(1:n_participants, length(dataset_sizes) + 1),
                             F = rep(lf, length(dataset_sizes) + 1),
                             t_er = rep(t_er, length(dataset_sizes) + 1),
                             A_correct = rep(a_c_mu, length(dataset_sizes) + 1),
                             A_incorrect = rep(a_f_mu, length(dataset_sizes) + 1),
                             A_correct_sd = rep(a_c_sd, length(dataset_sizes) + 1),
                             A_incorrect_sd = rep(a_f_sd, length(dataset_sizes) + 1)) %>%
  pivot_longer(F:A_incorrect_sd, names_to = "parameter", values_to = "original")

s4_par_comp <- left_join(s4_param_recov_long, s4_param_orig_long, by = c("n", "participant", "parameter"))
```

Comparison of original ACT-R parameters to recovered parameter estimates from the LBA:
```{r}
ggplot(s4_par_comp, aes(x = original, y = recovered, colour = parameter)) +
  facet_wrap(~ n, labeller = function(x) label_both(labels = x, sep = " = ")) +
  geom_abline(lty = 2) +
  geom_point() +
  # coord_equal() +
  labs(x = "Original parameter value",
         y = "Recovered parameter value",
         colour = "Parameter") +
  theme_paper
```

Difference between the original ACT-R parameters and the recovered parameter estimates from the LBA, as a function of datset size:
```{r}
s4_par_comp_error <- s4_par_comp %>%
  filter(parameter != "A_correct_sd") %>%
  mutate(parameter = factor(parameter, 
                            levels = c("A_correct", "A_incorrect", "A_incorrect_sd", "F", "t_er"),
                            labels  = c(expression(mu[c]), expression(mu[f]), expression(sigma[f]), expression(F), expression(t[er])))) %>%
  rowwise() %>%
  mutate(ae = abs(original - recovered)) %>%
  ungroup()

s4_par_comp_error_param <- s4_par_comp_error %>%
  group_by(n, parameter) %>%
  summarise(mae = mean(ae))

s4_par_comp_error_mean <- s4_par_comp_error %>%
  group_by(n, parameter) %>%
  summarise(mae = mean(ae)) %>%
  group_by(n) %>%
  summarise(mae = mean(mae))


p_par_comp_multi <- ggplot() +
  geom_rect(aes(xmin = exp(log(100) - .25), xmax = exp(log(100) + .25), ymin = 0, ymax = 2), colour = NA, fill = "grey90") +
  geom_jitter(data = s4_par_comp_error, aes(x = n, y = ae, colour = parameter), height = 0, width = .025, alpha = .05) +
  geom_line(data = s4_par_comp_error_param, aes(x = n, y = mae, group = parameter, colour = parameter), lty = 2) +
  geom_point(data = s4_par_comp_error_param, aes(x = n, y = mae, group = parameter, colour = parameter), alpha = .9) +
  geom_line(data = s4_par_comp_error_mean, aes(x = n, y = mae), lty = 2, lwd = rel(.85)) +
  geom_point(data = s4_par_comp_error_mean, aes(x = n, y = mae), size = rel(2)) +
  scale_x_log10(breaks = c(25, 50, 100, 250, 500, 1000, 2500, 5000, 10000, 25000, 50000),
                labels = c(25, 50, 100, 250, 500, "1K", "2.5K", "5K", "10K", "25K", "50K")) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_viridis_d() +
  coord_cartesian(ylim = c(0, 2)) +
  labs(x = "Number of trials",
       y = "Absolute error") +
  guides(colour = FALSE) +
  theme_paper +
  theme(axis.text.x = element_text(face = c("plain", "plain", "bold", "plain", "plain", "plain", "plain", "plain", "plain", "plain", "plain")))

p_par_comp_multi
```

Make version with colour legend for presentation:
```{r}
ggplot() +
  geom_rect(aes(xmin = exp(log(100) - .25), xmax = exp(log(100) + .25), ymin = 0, ymax = 2), colour = NA, fill = "grey90") +
  geom_jitter(data = s4_par_comp_error, aes(x = n, y = ae, colour = parameter), height = 0, width = .025, alpha = .05) +
  geom_line(data = s4_par_comp_error_param, aes(x = n, y = mae, group = parameter, colour = parameter), lty = 2) +
  geom_point(data = s4_par_comp_error_param, aes(x = n, y = mae, group = parameter, colour = parameter), alpha = .9) +
  geom_line(data = s4_par_comp_error_mean, aes(x = n, y = mae), lty = 2, lwd = rel(.85)) +
  geom_point(data = s4_par_comp_error_mean, aes(x = n, y = mae), size = rel(2)) +
  scale_x_log10(breaks = c(25, 50, 100, 250, 500, 1000, 2500, 5000, 10000, 25000, 50000),
                labels = c(25, 50, 100, 250, 500, "1K", "2.5K", "5K", "10K", "25K", "50K")) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_colour_viridis_d(labels = scales::parse_format()) +
  coord_cartesian(ylim = c(0, 2)) +
  labs(x = "Number of trials",
       y = "Absolute error",
       colour = NULL) +
  # guides(colour = FALSE) +
  theme_paper +
  theme(axis.text.x = element_text(face = c("plain", "plain", "bold", "plain", "plain", "plain", "plain", "plain", "plain", "plain", "plain")),
        legend.background = element_blank(),
        legend.direction = "horizontal",
        legend.position = "top",
        legend.justification = "right",
        legend.box.margin = unit(c(0, 0, -40, 0), "pt"))

ggsave(file.path("..", "output", "param-recov-multi-error.png"), width = 6, height = 3, dpi = 600)
```


Make the figure in the paper:
```{r}
plot_grid(p_par_comp, p_par_comp_multi,
          align = "hv",
          axis = "tblr",
          labels = c("A", "B"))

ggsave(file.path("..", "output", "param_recov_multi_combi.pdf"), width = 9, height = 4.5)
```




# Session info
```{r}
sessionInfo()
```

